### 2019-07-29 concept_dev.tex
template for ConDev

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}

%SetFonts

%SetFonts

\title{Concept Dev for COMP study}
\author{Sitaram Devarakonda \& George Chacko}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
\section{Motivation} Making discovery in a manner that would leverage existing resources and skills with focus on  the computer science field, which has experienced vigorous growth in the last twenty years. Citation analysis is a particular focus.

%\subsection{}
\end{document}